\normalfont \normalsize \textsc{university, school or department name} \\ [25pt] \horrule{0.5pt} \\[0.4cm] \huge Assignment Title \\ \horrule{2pt} \\[0.5cm]

### 2013-01-27 exercise3/Paper/paper.tex
added java stuff to paper

\author{Benjamin Kiesl \and Philipp Steinwender \and Robert Schäfer} % Your name

@@

@@

\section{Evaluation}
Our Data Sets were 
\begin{enumerate}
\item
    The 'Skin-NonSkin' Data Set, which is a collection of RGB-values and the 4th attribute states whether this is a skin color or not
\item
    A truncation of the iris data set, where we have removed the 4th column to adjust the data to three dimensions.
\item
    The 'Haberman's Survival' data set, which contains cases from a study on the survival of patients who had undergone surgery for breast cancer.
\end{enumerate}

>>>>

\author{Benjamin Kiesl \and Philipp Steinwender \and Robert Sch\"{a}fer} % Your name

@@

For this assignment we have two implementations: one in Java and a second one in R.

\subsection{kNN visualization in Java}

To visualize the decision boundaries and  classification results of weka's kNN implementation iBK a Java gui was implemented. We used Swing and plain Java. There is also a dataset generator for multiple gaussian cluster that can only be executed after editing the code and recompilation of the application.

To run the application, execute the \texttt{kNN/run.sh} script. There is also a version for windows that was not tested.

\begin{figure}[\textwidth]
    \begin{center}
\includegraphics[width=\textwidth]{knn-visualization-marker}
    \end{center}
\caption{Java GUI}
\label{fig:javagui}
\end{figure}

Figure \ref{fig:javagui} shows how the gui looks like and which controls there are for manipulating the parameters of the data set and the classifier. The drawing area on the left displays the data points on a two dimensional coordinate system. The coordinate values of the data set get scaled to fit this area. The two markers showing the numbers 1 and 2 were added after taking the screenshot and highlight the following features:

\begin{enumerate}
\item With these 4 blue data points can be described how we visualize the decision boundaries of knn. The blue point in the middle has a light grey bigger circle around it. Points that look like this are from the test set. Data point that do not have this light grey highlighting are data points from the training set. Each data point of the test set has \texttt{k} connectors to surrounding data points which are the k nearest neighbors of the point.
\item Here we can see a half red half green data point of the test set. It is connected with 2 green data points and a blue one. Therefore the classifier predicts its class as green. The predicted class can be seen in the right half of the data point. The left half shows which class the data point acutally has which is red. If theclassifier predicts the true value of a data point, both halfs have the same color.
\end{enumerate}

On the right side of the gui is a control panel. From top to bottom:

\begin{itemize}
\item The "open file" button can be used to read an arff file from disk. After selecting a file, it's content is loaded, a classifier is created and the result gets drawn. Below is a label that displays the name of the currently selected file.
\item The spinner with the label \texttt{k} can be used to modify the classifier's value of \texttt{k}.
\item The training-test-split can be modified with the spinner showing the number 70 whereby this means that we do a 70\% split and use 70\% for training and the rest for classification. Although kNN does not train a model, this parameter tells the classifier how much unclassified instances it should use.
\item The next three checkboxes can be used to view or hide specific visualization details. One can hide the data points of the training data, those of the test set and hide the connectors between the test data points and their nearest neighbors.
\item Then there is a legend telling the attribute names and which coloreach class is assigned to.
\end{itemize}

All controls have a ChangeListener that redraws the data points and the results immediately after a value got changed by the user.

When loading a new file, the data is read by weka's arff parser. The created \texttt{Instances} are split by the specified percentage using the \texttt{RemovePercentageFilter}. Before the spilt, the data is shuffled. Then the \texttt{iBK} classifier gets instantiated, configured and the data points are drawn.

With this visualization one can easily recognize how the classifier predicts the class of a new instance and how the values of k and the train-test-set-split can change runtime behaviour.

\subsection{kNN visualization in R}

@@

\subsubsection{Evaluation}
Our Data Sets were 
\begin{enumerate}
\item
    The 'Skin-NonSkin' Data Set, which is a collection of RGB-values and the 4th attribute states whether this is a skin color or not
\item
    A truncation of the iris data set, where we have removed the 4th column to adjust the data to three dimensions.
\item
    The 'Haberman's Survival' data set, which contains cases from a study on the survival of patients who had undergone surgery for breast cancer.
\end{enumerate}
